---
title: "Homework 2"
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
library(dplyr)
library(ggplot2)
```

# Problem 1

## Mr. Trash Wheel Dataset 
Reading the Mr. Trash Wheel dataset, and cleaning it by specifying the sheet, omitting non-data entries, changing variable names, omitting rows that do not include dumpster-specific data, and rounding the number of sports balls to the nearest integer. 

```{r}
trash_wheel = read_excel("./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", 
                         range = "A2:N535", 
                         "Mr. Trash Wheel") %>% 
  na.omit("Dumpster") %>% 
  janitor::clean_names() %>% 
  mutate(sports_balls = round(as.numeric(sports_balls)))

head(trash_wheel) %>% 
  knitr::kable()
```

The above table is a 6-row preview of the cleaned Mr. Trash Wheel dataset. 

## Precipitation Datasets
Reading and cleaning the precipitation data for 2018 and 2019 through omitting rows without precipitation data, adding a `year` variable, combining precipitation datasets and converting month to a character variable. 

```{r}
precipitation_2018 = read_excel("./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", 
                         range = "A2:B14", 
                         "2018 Precipitation") %>% 
  mutate(year = "2018")

precipitation_2019 = read_excel("./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", 
                         range = "A2:B14", 
                         "2019 Precipitation") %>% 
  mutate(year = "2019")

precipitation =
  bind_rows(precipitation_2018, precipitation_2019) %>% 
  janitor::clean_names() %>% 
  mutate(month = month.name[month])

head(precipitation) %>% 
  knitr::kable()
``` 

The trash wheel dataset has `r nrow(trash_wheel)` observations and `r ncol(trash_wheel)` variables, and key variables are `r names(trash_wheel)`. The median number of sports balls in a dumpster in 2019 was `r median(pull(filter(trash_wheel, year == 2019), sports_balls))` balls. 

The precipitation dataset has `r nrow(precipitation)` observations, and key variables are `r names(precipitation)`. The precipitation data ranges from `r min(pull(precipitation, total))` to `r max(pull(precipitation, total))`. The total precipitation in 2018 was `r sum(pull(precipitation_2018, Total))`. 

# Problem 2
## pols-month Dataset
Clean the data in pols-month.csv through breaking up the variable mon into year, month, and day, replacing month number with month name, creating a president variable. 
```{r}
pols_month = 
  read_csv("./data/fivethirtyeight_datasets/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, into = c("year", "month", "day")) %>% 
  mutate(year = as.integer(year), month = as.integer(month), day = as.integer(day)) %>% 
  mutate(month = month.name[as.numeric(month)]) %>% 
  mutate(president = ifelse(prez_gop == 1, "gop", "dem")) %>% 
  select(-"day", -"prez_gop", -"prez_dem") %>% 
  select(year, month, everything()) %>% 
  mutate(year = as.numeric(year)) 

head(pols_month) %>% 
  knitr::kable()
```

## snp Dataset
Cleaning the data in snp.csv by breaking up the variable date into year, month, and day, and replacing month number with month name.
```{r}
snp = 
  read_csv("./data/fivethirtyeight_datasets/snp.csv") %>% 
  janitor::clean_names() %>% 
  mutate(date, date = lubridate::mdy(date)) %>% 
  separate(date, into = c("year", "month", "day")) %>% 
  mutate(year = as.numeric(year)) %>% 
  mutate(year = ifelse(year > 2049, year - 100, year)) %>% 
  mutate(month = month.name[as.numeric(month)]) %>% 
  arrange(year, month) %>% 
  select(year, month, everything()) %>% 
  select(-"day")
  
head(snp) %>% 
  knitr::kable()
```
## Unemployment Data
Cleaning the unemployment data so that it has the same names of the key variables, and ensuring that key variables month and year take the same values.
```{r}
unemployment = 
  read_csv("./data/fivethirtyeight_datasets/unemployment.csv") %>% 
  pivot_longer(Jan:Dec,
               names_to = "month",
               values_to = "unemployment") %>% 
  janitor::clean_names() %>% 
  mutate(month = month.name[match(month, month.abb)]) %>% 
  arrange(year, month)%>% 
  mutate(year = as.numeric(year)) 

head(unemployment) %>% 
  knitr::kable()
```
## Joining Datasets
Merging snp into pols, and merging unemployment into the resulting dataframe.
```{r}
snp_pols = left_join(pols_month, snp, by = c("year", "month"))

merge_df = left_join(snp_pols, unemployment, by = c("year", "month"))

head(merge_df) %>% 
  knitr::kable()
```

The `pols-month.csv` dataset has `r nrow(pols_month)` observations and `r ncol(pols_month)` variables related to the number of national politicians who are democratic or republican. Variables include `r names(pols_month)`. The year ranges from `r min(pull(pols_month, year))` to `r max(pull(pols_month, year))`. The new `president` column indicates that whether the president was republican or democratic on the associated date. 

The `snp.csv` dataset has `r nrow(snp)` observations and `r ncol(snp)` variables related to Standard & Poor’s stock market index. Variables include `r names(snp)`. The year ranges from `r min(pull(snp, year))` to `r max(pull(snp, year))`. 

The `unemployment.csv` dataset has `r nrow(unemployment)` observations and `r ncol(unemployment)` variables of the percentages of unemployment. Variables include `r names(unemployment)`. The year ranges from `r min(pull(unemployment, year))` to `r max(pull(unemployment, year))`.

The resulting final dataset has `r nrow(merge_df)` observations and `r ncol(merge_df)` variables. Variables include `r names(merge_df)`. The year ranges from `r min(pull(merge_df, year))` to `r max(pull(merge_df, year))`. 

# Problem 3
## Tidying
Load and tidy the data by changing names in the ethnicity column, changing the case structure in the child's first name column, and removing duplicated rows. 

```{r}
names = 
  read_csv("./data/Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>% 
  mutate(ethnicity = str_to_lower(ethnicity), gender = str_to_lower(gender)) %>%
  mutate(ethnicity = recode(ethnicity, "asian and paci" = "asian and pacific islander", "black non hisp" = "black non hispanic", "white non hisp" = "white non hispanic")) %>% 
  mutate(childs_first_name = str_to_title(childs_first_name)) %>% 
  distinct()

head(names) %>% 
  knitr::kable()
```
## Rank in Popularity of the Name “Olivia” Over Time
To show the popularity of "Olivia" over time, we need to create a subtable from the original `names` dataset which only contains data relating to the child's first name "Olivia". We also need to make it into a wider dataframe which has rows for ethnicities and columns for year in order to see the trend of ranks over time. 
```{r}
olivia = 
  names %>% 
  filter(childs_first_name == "Olivia") %>% 
  select(-"gender", -"childs_first_name", -"count") %>% 
  arrange(year_of_birth) %>% 
  pivot_wider(
    names_from = "year_of_birth", 
    values_from = "rank")

head(olivia) %>% 
  knitr::kable()
```

## Most Popular Name Among Male Children Over Time

To show the most popular name among male children over time, we need to create a sub-dataframe from the original `names` dataset, and it should only contains data relating to all boy's names ranking 1 in each ethnicity group. We also need to make it into a wider dataframe which has rows for ethnicity and columns for year in order to see the trend of popular names over time. 

```{r}
popular =
  names %>% 
  filter(gender == "male", rank == "1") %>% 
  select(-"gender", -"count", -"rank") %>% 
  arrange(year_of_birth) %>% 
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "childs_first_name"
  )

head(popular) %>% 
  knitr::kable()
```

## Scatterplot for Male, White Non-hispanic Children Born in 2016

A sub-dataset is made by filtering for only male, 2016 and white non-hispanic.  

```{r}
male_white_non_hispanic_2016 = 
  names %>% 
  filter(gender == "male", year_of_birth == "2016", ethnicity == "white non hispanic")

male_white_non_hispanic_2016 %>% 
  ggplot(aes(x = rank, y = count)) +
  geom_point(alpha = .3) +
  ggtitle("Count vs. Rank for Male, White Non-Hispanic Children Born in 2016") +
  xlab("Rank in Popularity of that Name") +
  ylab("Number of Children with a Name") +
  theme_minimal()

ggsave("male_white_non_hispanic_2016.pdf")
```

