---
title: "Homework 2"
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
library(dplyr)
```

# Problem 1

Reading and cleaning the Mr. Trash Wheel dataset. 

```{r}
trash_wheel = read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
                         range = "A2:N408", 
                         1) %>% 
  na.omit("Dumpster") %>% 
  janitor::clean_names() %>% 
  mutate(sports_balls = round(as.numeric(sports_balls)))
```

Reading and cleaning the precipitation data for 2018. 

```{r}
precipitation_2018 = read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
                         range = "A2:B14", 
                         5) %>% 
  mutate(year = "2018")

precipitation_2019 = read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
                         range = "A2:B8", 
                         4) %>% 
  mutate(year = "2019")

precipitation =
  bind_rows(precipitation_2018, precipitation_2019) %>% 
  janitor::clean_names() %>% 
  mutate(month = month.name[month])
``` 

The trash wheel dataset has `r nrow(trash_wheel)` observations, and key variables are `r names(trash_wheel)`. The median number of sports balls in a dumpster in 2019 was `r median(pull(filter(trash_wheel, year == 2019), sports_balls))` . 

The precipitation dataset has `r nrow(precipitation)` observations, and key variables are `r names(precipitation)`. The total precipitation in 2018 was `r sum(pull(precipitation_2018, Total))`. 

# Problem 2

Reading the pols-month dataset. 


```{r}
pols_month = 
  read_csv("./data/fivethirtyeight_datasets/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, into = c("year", "month", "day")) %>% 
  mutate(month = month.name[as.numeric(month)]) %>% 
  select(-"day") %>% 
  pivot_longer(prez_gop:rep_dem,
               names_to = "president", 
               values_to = "values") %>% 
  filter(!president == "prez_dem", !president == "prez_gop")%>% 
  select(year, month, everything())%>% 
  mutate(month = substring(month, 0,3)) %>% 
  mutate(year = as.numeric(year))
```

Cleaning the snp dataset. 

```{r}
snp = 
  read_csv("./data/fivethirtyeight_datasets/snp.csv") %>% 
  janitor::clean_names() %>% 
  mutate(date, date = lubridate::mdy(date)) %>% 
  separate(date, into = c("year", "month", "day")) %>% 
  mutate(year = as.numeric(year)) %>% 
  mutate(year = ifelse(year>2049, year - 100, year)) %>% 
  mutate(month = month.name[as.numeric(month)]) %>% 
  arrange(year, month) %>% 
  select(year, month, everything()) %>% 
  mutate(month = substring(month, 0,3)) %>% 
  select(-"day")
  
```

Cleaning the unemployment data. 

```{r}
unemployment = 
  read_csv("./data/fivethirtyeight_datasets/unemployment.csv") %>% 
  pivot_longer(Jan:Dec,
               names_to = "month",
               values_to = "unemployment") %>% 
  janitor::clean_names() %>% 
  arrange(year, month)%>% 
  mutate(year = as.numeric(year))
  
```

Joining dataframes. 

```{r}
snp_pols = left_join(pols_month, snp, by = c("year", "month"))

merge_df = left_join(snp_pols, unemployment, by = c("year", "month"))
```


