---
title: "Homework 2"
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
library(dplyr)
```

# Problem 1

Reading and cleaning the Mr. Trash Wheel dataset. 

```{r}
trash_wheel = read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
                         range = "A2:N408", 
                         1) %>% 
  na.omit("Dumpster") %>% 
  janitor::clean_names() %>% 
  mutate(sports_balls = round(as.numeric(sports_balls)))
```

Reading and cleaning the precipitation data for 2018. 

```{r}
precipitation_2018 = read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
                         range = "A2:B14", 
                         5) %>% 
  mutate(year = "2018")

precipitation_2019 = read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
                         range = "A2:B8", 
                         4) %>% 
  mutate(year = "2019")

precipitation =
  bind_rows(precipitation_2018, precipitation_2019) %>% 
  janitor::clean_names() %>% 
  mutate(month = month.name[month])
``` 

Write a paragraph about these data; you are encouraged to use inline R. Be sure to note the number of observations in both resulting datasets, and give examples of key variables. For available data, what was the total precipitation in 2018? What was the median number of sports balls in a dumpster in 2019?

The trash wheel dataset has `r nrow(trash_wheel)` observations, and key variables are `r names(trash_wheel)`. The median number of sports balls in a dumpster in 2019 was `r median(pull(trash_wheel, sports_balls), year = "2019")` 

The precipitation dataset has `r nrow(precipitation)` observations, and key variables are `r names(precipitation)`. The total precipitation in 2018 was `r sum(pull(precipitation_2018, Total))`. 

# Problem 2

Reading the pols-month dataset. 


