---
title: "Homework 2"
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
library(dplyr)
library(ggplot2)
```

# Problem 1

Reading and cleaning the Mr. Trash Wheel dataset. 

```{r}
trash_wheel = read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
                         range = "A2:N408", 
                         1) %>% 
  na.omit("Dumpster") %>% 
  janitor::clean_names() %>% 
  mutate(sports_balls = round(as.numeric(sports_balls)))
```

Reading and cleaning the precipitation data for 2018. 

```{r}
precipitation_2018 = read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
                         range = "A2:B14", 
                         5) %>% 
  mutate(year = "2018")

precipitation_2019 = read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
                         range = "A2:B8", 
                         4) %>% 
  mutate(year = "2019")

precipitation =
  bind_rows(precipitation_2018, precipitation_2019) %>% 
  janitor::clean_names() %>% 
  mutate(month = month.name[month])
``` 

The trash wheel dataset has `r nrow(trash_wheel)` observations, and key variables are `r names(trash_wheel)`. The median number of sports balls in a dumpster in 2019 was `r median(pull(filter(trash_wheel, year == 2019), sports_balls))` . 

The precipitation dataset has `r nrow(precipitation)` observations, and key variables are `r names(precipitation)`. The total precipitation in 2018 was `r sum(pull(precipitation_2018, Total))`. 

# Problem 2

Reading the pols-month dataset. 
‘

```{r}
pols_month = 
  read_csv("./data/fivethirtyeight_datasets/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, into = c("year", "month", "day")) %>% 
  mutate(year = as.integer(year), month = as.integer(month), day = as.integer(day)) %>% 
  mutate(month = month.name[as.numeric(month)]) %>% 
  mutate(president = ifelse(prez_gop == 1, "gop", "dem")) %>% 
  select(-"day", -"prez_gop", -"prez_dem") %>% 
  select(year, month, everything()) %>% 
  mutate(year = as.numeric(year))
```

Cleaning the snp dataset. 

```{r}
snp = 
  read_csv("./data/fivethirtyeight_datasets/snp.csv") %>% 
  janitor::clean_names() %>% 
  mutate(date, date = lubridate::mdy(date)) %>% 
  separate(date, into = c("year", "month", "day")) %>% 
  mutate(year = as.numeric(year)) %>% 
  mutate(year = ifelse(year > 2049, year - 100, year)) %>% 
  mutate(month = month.name[as.numeric(month)]) %>% 
  arrange(year, month) %>% 
  select(year, month, everything()) %>% 
  select(-"day")
  
```

Cleaning the unemployment data. 

```{r}
unemployment = 
  read_csv("./data/fivethirtyeight_datasets/unemployment.csv") %>% 
  pivot_longer(Jan:Dec,
               names_to = "month",
               values_to = "unemployment") %>% 
  janitor::clean_names() %>% 
  mutate(month = month.name[match(month, month.abb)]) %>% 
  arrange(year, month)%>% 
  mutate(year = as.numeric(year)) 
  
```

Joining dataframes. 

```{r}
snp_pols = left_join(pols_month, snp, by = c("year", "month"))

merge_df = left_join(snp_pols, unemployment, by = c("year", "month"))
```

# Problem 3



```{r}
names = 
  read_csv("./data/Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>% 
  mutate(ethnicity = str_to_lower(ethnicity), gender = str_to_lower(gender)) %>%
  mutate(ethnicity = recode(ethnicity, "asian and paci" = "asian and pacific islander", "black non hisp" = "black non hispanic", "white non hisp" = "white non hispanic")) %>% 
  mutate(childs_first_name = str_to_title(childs_first_name)) %>% 
  distinct()
```

```{r}
olivia = 
  names %>% 
  filter(childs_first_name == "Olivia") %>% 
  select(-"gender", -"childs_first_name", -"count") %>% 
  arrange(year_of_birth) %>% 
  pivot_wider(
    names_from = "year_of_birth", 
    values_from = "rank"
  )

```

```{r}
popular =
  names %>% 
  filter(gender == "male", rank == "1") %>% 
  select(-"gender", -"count", -"rank") %>% 
  arrange(year_of_birth) %>% 
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "childs_first_name"
  )
```



```{r}
male_white_non_hispanic_2016 = 
  names %>% 
  filter(gender == "male", year_of_birth == "2016", ethnicity == "white non hispanic") %>% 
  ggplot(aes(x = rank, y = count)) +
  geom_point()
```

